Supplementary Appendix

### 2023-06-29 manuscript/manuscript.qmd
first complete draft of manuscript and supplement (before co-author review)

---
title: "The FAIRification of pharmacoepidemiological research: A practical introduction to reproducible analytic workflows using git and R"

@@

[**Word count:**]{.underline} xxxx words

@@

[**Supplementary material:**]{.underline}

[**Short running title**]{.underline}: FAIRification of pharmacoepidemiological research

@@

[**Acknowledgments:**]{.underline} The authors would like to thank ...

@@

**Purpose**

**Study Design and Settings**

**Results**

**Conclusions**

@@

Transparency and reproducibility in conducting healthcare database studies in pharmacoepidemiology are critical scientific requirements for meaningful research. While many advances have been made in the documentation and reporting of study protocols and results[@wang2021; @schneeweiss2019; @berger2017; @wang2022], the transparency around the actual implementation and analytic steps is still significantly lacking, especially when analyses are implemented via line programming steps in statistical programming languages like R[@RCore2023] or Python. Although in other quantitative disciplines such as computational biology[@almugbel2017; @mammoliti2021; @russell2018] or health informatics[@bakken2019], the sharing of programming code is rather the rule than the exception, this practice is not as common in the (pharmaco-)epidemiological community.

Version control systems (VCS), such as *Git*, can thereby provide easy and powerful tools to keep track of the versioning of important files and documents such as protocols, analytical code, tables and figures. Thereby, it can extend the principles of *FAIR* *data*[@garcía-closas2023] to the implementation of pharmacoepidemiological study workflows by making all components of a study **F**indable, **A**ccessible, **I**nteroperable, and **R**eproducible (**FAIR**).

VCS enable researchers to track and resolve errors, collaborate with peers, and share resources instantaneously, for example, upon submission or publication of a manuscript. By design, VCS workflows automatically empower users to comment, track and compare changes made to files and hence, increase the ability to comprehend the evolution of a project over time while maintaining a complete audit trail of all documents.[@ram2013] Without VCS, this is often naturally done by assigning uninformative file names such as "*final_analysis_1_rev_more_changes.R*" which is bad practice as this is highly error prone, lacking transparency for collaborating project members and outside personnel, and changes are difficult to track and reconcile. Particularly in disciplines like pharmacoepidemiology, where analyses can have far reaching impact on healthcare decisions, analytical code needs to be accessible and reproducible by anyone, not only the analyst who originally carried out the analysis.

To increase the adoption of transparent and reproducible workflows in real-world evidence (RWE) studies using VCS, this tutorial aims to give a practical introduction tailored for pharmacoepidemiologists on how to set up, structure, and implement workflows using *Git*, which is the most frequently used VCS to date.[@blischak2016] We will first give a brief introduction to Git and how it can be used for collaboration and dissemination of study results using project repositories. Then we will provide a technical step-by-step guidance on how to integrate Git in an analytic workflow. We will additionally discuss and showcase important aspects of reproducibillity using the R open-source programming language, although the basic principles are applicable to any major coding language.

@@

# A step-by-step introduction on how to use Git in a pharmacoepidemiology research study

@@

To check if the installation succeeded and show the available version, the following prompt in the CLI can be used.

@@

An alternative to creating a new repository locally is through creating a new project in the remote repository (in this case GitHub) and then creating a local copy, a process which is usually referred to as `cloning`.

To do so, users will need to log into their remote repository account and navigate to the section `New repository` (**Supplementary Figure 1**). This will take a user to a form to populate basic information of the new project such as the repository name (`rwe_study`) and choose basic setting such as if the repository should be private or public and if a `README.md` (see @sec-repository-structure) and `.gitignore` (see @sec-stage-and-commit-changes) file should be automatically created (**Supplementary Figure 2**).

@@

At this step, we now can populate the repository and start working on study-relevant documents. To enhance clarity, it is recommended to use a logical structure and informative and transparent sub-directory and file names (a minimal example is illustrated in @fig-structure). To document the structure and location of files, we can include this and other critical information, that a peer or outside personnel should know concerning the project, in the `README.md` file. This file can be auto-generated when a project is initialized in a remote repository (see prior step in @sec-gui-initialization-and-git-clone) or can be added manually aftwerwards by creating and adding a text file with the name `README.md` in the root directory of the repository.

@@

Every time a user makes a commit to create a snapshot of the work, `Git` creates a unique hash (a 40-character string created by a **simple hashing algorithm** \[SHA\]) which can be seen as an "ID" of the snapshot which enables user to comprehend every change made and revert back to any snapshot of the repository. @fig-stage-commit-RStudio-gui illustrates the equivalent step by using the RStudio GUI.

If any file(s) in the study repository should not be tracked and hence not be synchronized with the remote repository (e.g., because they include confidential information that should not be visible to others like database credentials), these files can be specified in the `.gitignore` file which needs to be located in the root directory of the repository. To create a `.gitignore` file, either auto-generate it upon initialization of the repository on GitHub (**Supplementary Figure 2**) or manually add a text file called `.gitignore` to the root directory.

@@

To synchronize the local changes made to our files, the next step involves synchronizing our commits with the remote repository on GitHub via the `push` command.

@@

If there is no remote established yet, this can be configured using the following commands.

```{bash}
git remote add origin https://github.com/USERNAME/rwe_study.git
git branch -M main
git push -u origin main
```

This step is not needed if the intention of a researcher is just to keep track of local changes and there is no wish to share code, collaborate with others or work across different systems. However, most of the times researchers intentionally use `Git` to collaborate on a project and remote repositories such as GitHub as this can improve collaborative work significantly (@fig-collaborate).

@@

Occasionally it can occur that two collaborators made conflicting changes to the same file or even to the same line of code. While Git is generally good at automatically integrating (`merging`) new changes, Git will notify the user to resolve a major conflict by editing the conflicting file if Git cannot automatically determine what is correct ((e.g., when one collaborator has made changes to a line of code and the other completely deleted it).

@@

One of the main motivations to adhere to this workflow is that in return we get rewarded with a detailed audit trail of changes made to our files. Probably, the best way to visualize and track changes is by browsing the `commit history` on the remote repository. @fig-GitHub-gui_changes illustrates this for the example commit we have previously performed in @sec-stage-and-commit-changes. A similar view for uncommitted changes is also provided in the RStudio GUI as displayed in @fig-stage-commit-RStudio-gui.

@@

A complete summary of committed changes with details on SHA hash, author, email, time and commit messages can be accessed via

@@

@@

A further step to improve reproducibility when using the R programming language is to ensure that anyone intending to reproduce analytic results from a `Git` repository is able to work off the same R package versions and dependencies without having to perform manual installations or path adjustments. These requirements are nothing exclusive to R, but apply to any type of programming language.

@@

One way to achieve this is through self-contained computing environments via Docker images which are isolated and standardized environments which include everything needed to run and deploy applications and code regardless of a user's local operating system.[@boettiger2017a; @nüst2020] Another more lightweight, yet effective way is by using the `renv` R package[@renv], which helps creating **re**producible **env**ironments and manages R package dependencies for a given project. Through a lockfile, the `renv` package keeps track of the underlying version of R that is being used as well as all R packages, their versions and the installation source (e.g., <https://cran.r-project.org/>). To synchronize the R environment with that of a collaborator, e.g., after cloning a repository, all that needs to be done is running the following command in R.

@@

Appending the session info to the end of an R script can provide further context.

@@

```{r}
icd_codes <- read.csv("C:\Users\Username\Projects\rwe_study\protocol\icd_codes.csv")
```

The use of absolute paths will make it impossible for anyone to run the code without having to manually adapt every single path throughout a repository which is tedious and time consuming. A better solution to this is the use of relative paths. A popular R package to implement this is the `here` package which enables file referencing in project-oriented workflows. The main function "`here"` creates paths relative to the top-level of the project directory (in our example `rwe_study`). Applied to our example project, the following command would automatically reference the correct path and could be used without modification by any user re-running this script.

@@

`Git` is a powerful tool that perfectly integrates into workflows of quantitative disciplines which have to handle analytic code. It further provides features that make it convenient for project-oriented setups and to collaborate at scale. This tutorial aimed to give an introduction to essential concepts that make it useful for pharmacoepidemiology research, but is by no means comprehensive as `Git` provides a vast amount of additional features which were not addressed in this manuscript. To get familiar with more advanced `Git` usage, we compiled a non-exhaustive list of further readings in the **Supplementary Material**.

`Git` was primarily designed to work with source code files and works best for plain, text-based documents such programming code, rich text format (RTF), Markdown or LaTex files. There are limitations when it comes to tracking binary files like Word (.docx) and Excel (.xlsx) as these need to be interpreted by a program or a hardware processor[@blischak2016]. While it can be still useful to track versions of these file types, it is important to recognize that modifications in such documents can't be easily inspected using Rstudio or GitHub.

@@

Despite its potential initial complexities to get acquanited with the `Git` workflow, distributed VCS and code sharing provide unique advantages:

-   Full audit trail of which changes were made by whom, when and why

-   Even if the code is not shared, Git makes it easier for an individual to go back to the code

-   Git can serve as internal and external repository for code and made accessible to everyone, e.g. along with manuscript without copy-paste into word files or similar

-   It is free and open-source and reliable

-   Distributed VCS fully mirror entire repositories and hence every local clone is a full backup of the data

@@

```{r}
#| label: fig-RStudio-gui
#| fig-cap: "Clone a remote repository using the RStudio graphical user interface."
#| eval: true
#| echo: false
#| out-width: 50%
knitr::include_graphics(here::here("figures", "Figure_1_RStudio_init.png"))
```

@@

```{r}
{r}
#| label: fig-GitHub-gui_changes
#| fig-cap: "Remote repositories like GitHub provide tools to visually track changes made to analytical code and corresponding output such as figures. This commit history illustrates details about the commit (e.g., commit SHA hash and message), a side-by-side comparison of the previous (upper left) and modified (upper right) version of the figure and the corresponding changes made to the R code (bottom)."
#| out-width: 105%

include_graphics(here("figures", "Figure_6_commit_Github_gui.png"))
```

>>>>

---
title: "The FAIRification of research in real-world evidence: A practical introduction to reproducible analytic workflows using Git and R"

@@

[**Word count:**]{.underline} 3,028 words

@@

[**Supplementary material:**]{.underline} Supplementary Appendix and Supplementary Figures (pdf)

[**Short running title**]{.underline}: FAIRification of research in real-world evidence

@@

[**Acknowledgments:**]{.underline} None.

@@

Transparency and reproducibility are major prerequisites for meaningful real-world evidence (RWE) studies. Many advances have been made in the documentation and reporting of study protocols and results, but the version control and sharing of analytic code in RWE is not as established yet as in other quantitative disciplines like computational biology and health informatics. In this practical tutorial, we aim to give an introduction to distributed version control systems (VCS) tailored towards the FAIR (**F**indable, **A**ccessible, **I**nteroperable and **R**eproducible) implementation of RWE studies. To ease adoption, we provide detailed step-by-step instructions with practical examples on how the Git VCS and R programming language can be implemented into RWE workflows to facilitate reproducible analyses. We further discuss and showcase how these tools can be used to track changes, collaborate, disseminate and archive RWE studies through dedicated project repositories maintaining a complete audit trail of all relevant study documents.

@@

Transparency and reproducibility in conducting real-world evidence (RWE) studies are critical scientific requirements for meaningful research. While many advances have been made in the documentation and reporting of study protocols and results[@wang2021; @schneeweiss2019; @berger2017; @wang2022], the transparency around the actual implementation and analytic steps is still significantly lacking, especially when analyses are implemented via line coding in statistical programming languages like R[@RCore2023] or Python. Although in other quantitative disciplines such as computational biology[@almugbel2017; @mammoliti2021; @russell2018] or health informatics[@bakken2019], the sharing of code is rather the rule than the exception, this practice is not as common in the (pharmaco-) epidemiological community.

Version control systems (VCS), such as *Git*, can thereby provide easy and powerful tools to keep track of the versioning of important files and documents such as protocols, analytic code, tables and figures. Thereby, it can extend the principles of *FAIR* *data*[@garcía-closas2023] to the implementation of pharmacoepidemiological study workflows by making all components of a study **F**indable, **A**ccessible, **I**nteroperable, and **R**eproducible (**FAIR**).

VCS enable researchers to track and resolve errors, collaborate with peers, and share resources instantaneously, for example, upon publication of a manuscript. By design, VCS workflows automatically empower users to comment, track and compare changes made to files and hence, increase the ability to comprehend the evolution of a project over time while maintaining a complete audit trail of all documents.[@ram2013] Without VCS, this is often naturally done by assigning uninformative file names such as "*final_analysis_1_rev_more_changes.R*" which is bad practice as this is highly error prone, lacking transparency for collaborators, and changes are difficult to track and reconcile. Particularly in disciplines like pharmacoepidemiology, where analyses can have far reaching impact on healthcare decisions, analytic code needs to be accessible and reproducible.

To increase the adoption of transparent and reproducible workflows in RWE studies using VCS, this tutorial aims to give a practical introduction for pharmacoepidemiologists on how to set up, structure, and implement workflows using *Git*, which is the most popular VCS to date.[@blischak2016] We will first give a brief introduction to Git and its usage for collaboration and dissemination of study results through project repositories. Then we will provide a technical step-by-step guidance on how to integrate Git in analytic RWE workflows. We will additionally discuss and showcase important aspects of reproducibillity using the R open-source programming language, although the basic principles are applicable to any major coding language.

@@

# A step-by-step introduction on how to use Git in a RWE study

@@

To check if the installation succeeded, the following prompt in the CLI can be used, which will output the available version.

@@

An alternative to creating a new repository locally is through initiating a new project in the remote repository (in this case GitHub) and then creating a local copy, a process which is usually referred to as `cloning`.

To do so, users will need to log into their remote repository account and navigate to the section `New repository` (**Supplementary Figure 1**). This will take a user to a form to populate basic information of the new project such as the repository name (`rwe_study`) and choose basic settings such as if the repository should be private or public and if `README.md` (see @sec-repository-structure) and `.gitignore` (see @sec-stage-and-commit-changes) files should be automatically created (**Supplementary Figure 2**).

@@

At this step, we now can populate the repository and start working on study-relevant documents and analyses. To enhance clarity, it is recommended to use a logical structure and an informative nomenclature for sub-directories and file names (a minimal example is illustrated in @fig-structure). To document the structure and location of files, we can include this, and other critical information about the project, in the `README.md` file. This file can be auto-generated when a project is initialized through a remote repository (see prior step in @sec-gui-initialization-and-git-clone) or can be manually added aftwerwards by creating a text file with the name `README.md` in the root directory of the repository.

@@

Every time a user makes a commit to create a snapshot of the work, `Git` creates a unique hash (a 40-character string created by a **simple hashing algorithm** \[SHA\]) which can be seen as an "ID" of the snapshot which enables users to comprehend every change made and revert back to any snapshot of the repository. @fig-stage-commit-RStudio-gui illustrates the equivalent steps in RStudio GUI.

If any file(s) in the study repository should not be tracked and hence not be synchronized with the remote repository (e.g., because they include confidential information that should not be visible to others like database credentials), these can be specified in the `.gitignore` file which is typically located in the root directory of the repository. To create a `.gitignore` file, a user can either auto-generate it upon initialization of the repository on GitHub (**Supplementary Figure 2**) or manually add a text file called `.gitignore` to the root directory.

@@

To synchronize the local changes, the next step involves the upload of our commits to the remote repository on GitHub via the `push` command.

@@

If there is no remote established yet, this can be retrospectively configured using the following commands.

```{bash}
git remote add origin https://github.com/USER/rwe_study.git
git branch -M main
git push -u origin main
```

This step is not needed if the intention of a researcher is just to keep track of local changes and there is no wish to share code, collaborate with others or work across different systems. However, most of the times researchers use `Git` for the purpose of collaborating on a project as depicted in @fig-collaborate.

@@

Occasionally it can occur that two collaborators made conflicting changes to the same file or even to the same line of code. While Git is generally good at automatically integrating (`merging`) new changes, Git will notify the user to resolve a major conflict by editing the conflicting file if Git cannot automatically determine what is correct (e.g., if one collaborator has made changes to a line of code and the other completely deleted it).

@@

One of the main motivations to adhere to this workflow is that in return we get rewarded with a detailed audit trail of changes made to our study files. Probably the best way to visualize and track changes is by browsing the `commit history` on the remote repository. @fig-GitHub-gui_changes illustrates this for the example commit we have previously performed in @sec-stage-and-commit-changes. A similar view for uncommitted changes in the RStudio GUI is also provided in @fig-stage-commit-RStudio-gui.

@@

A complete summary of committed changes with details on the SHA hash, author, email, time and commit messages can be accessed via

@@

While there are many more useful `Git` features such as `branching` (see **Supplementary Material**), the workflow of staging, committing and pushing/pulling (@fig-workflow) may already suffice for the majority of pharmacoepidemiologists.

@@

A further step to improve reproducibility when using the R programming language is to ensure that anyone, intending to reproduce analytic results from a `Git` repository, is able to work off the same R package versions and dependencies without having to perform manual installations or path adjustments. These requirements are nothing exclusive to R, but apply to any type of programming language.

@@

One way to achieve this is through self-contained computing environments via Docker images which are isolated and standardized environments which include everything needed to run and deploy applications and code regardless of a user's local operating system.[@boettiger2017a; @nüst2020] Another more lightweight, yet effective way is by using the `renv` R package[@renv], which helps creating **re**producible **env**ironments and manages R package dependencies for a given project. Through a lockfile, the `renv` package keeps track of the underlying version of R that is being used as well as all R packages, their versions and the installation source (e.g., <https://cran.r-project.org/>). To synchronize the R environment with that of a collaborator, e.g., after cloning a repository, the only step to necessary is running the following command in R.

@@

Appending the session info to the end of an R script can provide further context about the dependencies and versions of an analytic script.

@@

```{r}
icd_codes <- read.csv('C:/User/Projects/rwe_study/protocol/icd_codes.csv')
```

The use of absolute paths will make it impossible for any collaborator to run the code without having to manually adapt every single path throughout a repository which is tedious and time consuming. A better solution to this is the use of relative paths. A popular R package to implement this is the `here` package which enables file referencing in project-oriented workflows. The main function "`here"` creates paths relative to the top-level of the project directory (in our example the `rwe_study` directory). Applied to our example project, the following command would automatically reference the correct path and could be used without modification by any user re-running this script.

@@

`Git` is a powerful tool that perfectly integrates into workflows of quantitative disciplines handling analytic code. It further provides features that make it convenient for project-oriented setups and collaboration at scale. This tutorial is by no means comprehensive as `Git` provides a vast amount of additional features which were not addressed in this manuscript. To get familiar with more advanced `Git` usage, we compiled a non-exhaustive list of further suggested readings and resources in the **Supplementary Material**.

`Git` was primarily designed to work with source code files and works best for plain, text-based documents such as programming code, rich text format (RTF), Markdown or LaTex files. There are limitations when it comes to tracking binary files like Word (.docx) and Excel (.xlsx) as these need to be interpreted by a program or a hardware processor.[@blischak2016] While it can be still useful to track versions of these file types using `Git`, it is important to recognize that modifications in such documents can't be easily inspected using RStudio or GitHub.

@@

Despite the potential initial complexities to get acquanited with the `Git` workflow, distributed VCS and code sharing provide unique advantages, such as a full audit trail of which changes were made by whom, when and why. Git makes it easier for users to go back to any snapshot and resolve errors more efficiently. Further, remote repositories are ideal platforms to deploy and archive scripts and documents, facilitate collaboration and make it very convenient to share code. Although data sharing is unfortunately often prohibited due to privacy and legal reasons, many databases have been mapped to common data models like OMOP[@ohdsi2019book] which makes shared code interoperable even without direct access to the raw data. Finally, code sharing can enable researchers to learn from each other, build trust[@orsini2020], expedite the uptake of novel methodologies and knowledge exchange and decrease unneccesary redundancies through sharing of code libraries, functions and algorithms.

::: {#refs}
:::

@@

```{r}
#| label: fig-RStudio-gui
#| fig-cap: "Steps to clone a remote repository using the RStudio graphical user interface."
#| eval: true
#| echo: false
#| out-width: 50%
knitr::include_graphics(here::here("figures", "Figure_1_RStudio_init.png"))
```

@@

```{r}
#| label: fig-GitHub-gui_changes
#| fig-cap: "Remote repositories like GitHub provide tools to visually track changes made to analytical code and corresponding output such as figures. This commit history illustrates details about the commit (e.g., commit SHA hash and message), a side-by-side comparison of the previous (upper left) and modified (upper right) version of the figure and the corresponding changes made to the R code (bottom)."
#| eval: true
#| echo: false
#| out-width: 105%

knitr::include_graphics(here("figures", "Figure_6_commit_Github_gui.png"))
```

### 2023-06-29 manuscript/supplement.qmd
first complete draft of manuscript and supplement (before co-author review)

---
title: "Supplementary Appendix"
subtitle: "The FAIRification of pharmacoepidemiological research: A practical introduction to reproducible analytical workflows using git and R"
code-fold: true
echo: false
number-sections: true

format: 
  pdf:
    fig-cap-location: top
    fig-align: center
    crossref:
      fig-title: "**Supplementary Figure**"
    toc: true
    toc-depth: 2
    number-sections: false
    colorlinks: true

@@

@@

include_graphics(here("manuscript/figures", "SFigure_1a_new_repo.png"))
```

@@

include_graphics(here("manuscript/figures", "SFigure_2_repo_info.png"))
```

```{r}
#| label: fig-GitHub-gui_c
#| fig-cap: "Clone a GitHub repository using the HTTPS link and the git clone command"
#| out-width: 115%

include_graphics(here("manuscript/figures", "SFigure_3_clone.png"))
```

@@

# References {.unnumbered}

>>>>

---
title: "Supplementary Appendix"
subtitle: "The FAIRification of research in real-world evidence: A practical introduction to reproducible analytic workflows using Git and R"
code-fold: true
echo: false
number-sections: true

format: 
  pdf:
    fig-cap-location: top
    fig-align: center
    crossref:
      fig-title: "Supplementary Figure"
    toc: true
    toc-depth: 2
    number-sections: false
    colorlinks: true

@@

# Supplementary material

## Branching

Sometimes it is necessary to implement changes to a document or code without wanting to interfere with a current \"`production`\" version of a document or code. However, a researcher may still want to be able to \"`merge`\" back potential changes, e.g. to integrate new analyses into a script or to develop a new set of features or functions to an application. To do this in a safe, transparent and efficient manner, `Git` provides the option to **branch** our repository. The default branch of any repository is the `main` branch.

Branching is a concept that is highly adapted in software development where the `main` branch reflects the software version that is deployed for public use (i.e., the production version) and a `development` branch that is the version that implements new features (i.e., the beta version). Hence, `branches` are used in order to have the flexibility to develop new features without jeopardizing the current `main` and `development` versions.

Branching can have two very useful use cases for adaption in real-world evidence (RWE) studies. First, new analyses or improvements of existing analyses can be implemented without having to "mess" with existing code which works perfectly. Once a new analysis or visualization of results was established in a new `named branch` and works without errors, this branch can be `merged` back to the `main` branch (as depicted in the illustration below).

```{r}
#| label: branching
#| out-width: 100%

include_graphics(here("figures", "SFigure_x_branching.png"))
```

Second, branches can be re-purposed as backup versions of the repository at project landmarks. Examples could range from the first round of review among co-authors to the initial submission of a manuscript draft to a journal, corresponding rounds of revisions and the final version. This way it can be ensured that also prior versions of analyses can be reproduced without problems.

To create a new branch and switch to it from the main branch, you can run the `git checkout` command with the -b switch. Let's say we call the new branch `lasso_variable_selection`.

```{bash, eval=FALSE, echo=TRUE}
git checkout -b lasso_variable_selection
```

This is shorthand for:

```{bash, eval=FALSE, echo=TRUE}
git branch lasso_variable_selection
git checkout lasso_variable_selection
```

It is possible to switch between multiple branches back and forth and it's important to keep in mind that changes need to be committed before switch to the new branch branch. A user can e.g. switch back to the master branch by running the following command:

```{bash, eval=FALSE, echo=TRUE}
git checkout master
```

Suppose we have decided that the `lasso_variable_selection` approach led to a better balancing of the two new user cohorts after propensity score matching and we are ready to merge the implemented changes back into our main branch. This can be achieved by checking out the branch you wish to merge into (in this case `main` and then run the git merge command:

```{bash, eval=FALSE, echo=TRUE}
git checkout main
git merge lasso_variable_selection
```

If the branch is merged and there is no desire anymore to use this branch any longer, the `lasso_variable_selection` branch can be deleted with the following command:

```{bash, eval=FALSE, echo=TRUE}
git branch -d lasso_variable_selection
```

These steps can also easily be done on remote repositories like GitHub or GitLab. The only difference to keep in mind is that GitHub counterintuitively calls the procedure of merging two branches `pull request` while GitLab refers to it as `merge request`.

In case `Git` recognizes any clashes or incompatibilities of the new code with the existing code it will notify the user and give us the chance to reconcile inconsitencies.

## Advanced readings on `Git` and other resources

-   Main `Git` website to download and learn more about `Git:` <https://git-scm.com>

-   [Pro Git](https://git-scm.com/book/en/v2) by Scott Chacon and Ben Straub.

-   Cheatsheet on most important `Git` commands: <https://education.github.com/git-cheat-sheet-education.pdf>

-   GitHub Desktop App: <https://desktop.github.com/>

-   Git manual by *Atlassian* (provider of the `Bitbucket` platform): <https://www.atlassian.com/git>

-   Coursera courses:

    -   [Introduction to Git and GitHub](https://www.coursera.org/learn/introduction-git-github) (Google)

    -   [Version Control with Git](https://www.coursera.org/learn/version-control-with-git) (Atlassian)

    -   [Getting Started with Git and GitHub](https://www.coursera.org/learn/getting-started-with-git-and-github) (IBM)

    -   [Making Data Science Work for Clinical Reporting](https://www.coursera.org/learn/making-data-science-work-for-clinical-reporting?=) (Genentech)

-   Quarto (open-source scientific and technical publishing system for reproducible research): <https://quarto.org>

-   Choosing a suitable license for a remote repository: Public repositories need to be licensed to be truly open source. If no license is provided, the work is by default under exclusive copyright. The website <https://choosealicense.com> provides an excellent overview of the different types of licenses.

-   Useful Git repositories for observational healthcare research:

    -   Division of Pharmacoepidemiology and Pharmacoeconomics: [https://gitlab-scm.partners.org/drugepi](https://gitlab-scm.partners.org/drugepi/harper){.uri}

    -   Observational Health Data Sciences and Informatics (OHDSI): <https://github.com/OHDSI>

    -   Darwin EU® Private Development Repository: <https://github.com/darwin-eu-dev>

    -   Pharmaverse (Network of developers of curated open source R packages for clinical reporting): [https://pharmaverse.org](https://pharmaverse.orghttps://pharmaverse.org) & <https://github.com/pharmaverse/>

@@

include_graphics(here("figures", "SFigure_1a_new_repo.png"))
```

@@

include_graphics(here("figures", "SFigure_2_repo_info.png"))
```

```{r}
#| label: fig-GitHub-gui_c
#| fig-cap: "Clone a GitHub repository using the HTTPS link and the git clone command"
#| out-width: 105%
include_graphics(here("figures", "SFigure_3_clone.png"))
```

@@

```{r}
#| label: fig-GitHub-gui_changes
#| eval: false
#| fig-cap: "Remote repositories like GitHub provide tools to visually track changes made to analytical code and corresponding output such as figures. This commit history illustrates details about the commit (e.g., commit SHA hash and message), a side-by-side comparison of the previous (upper left) and modified (upper right) version of the figure and the corresponding changes made to the R code (bottom)."
#| out-width: 105%

include_graphics(here("figures", "SFigure_4_commit_Github_gui.png"))
```
